---
title: "MUR SST Anomaly — Florida Keys"
subtitle: "Daily sea surface temperature anomalies from NOAA CoastWatch ERDDAP"
format: html
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(httr2)
library(htmltools)
library(jsonlite)

erddap_base <- "https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41anom1day"
dataset_id <- "jplMURSST41anom1day"
n_days <- 5

# Florida Keys and adjacent South Florida shelf
lat_min <- 24.3
lat_max <- 26.5
lon_min <- -83.0
lon_max <- -79.5

fetch_times <- function() {
  resp <- request(paste0(erddap_base, ".json?time")) |>
    req_perform()

  rows <- resp_body_json(resp)$table$rows
  vapply(rows, function(row) row[[1]], character(1))
}

mur_image_url <- function(time_iso) {
  sprintf(
    paste0(
      "%s.smallPng?sstAnom[(%s)][(%.1f):(%.1f)][(%.1f):(%.1f)]",
      "&.draw=surface",
      "&.vars=longitude|latitude|sstAnom",
      "&.colorBar=|RedBlue| | |-3|3|",
      "&.land=over"
    ),
    erddap_base,
    time_iso,
    lat_min,
    lat_max,
    lon_min,
    lon_max
  )
}

format_time_label <- function(time_iso) {
  format(
    as.POSIXct(time_iso, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    "%Y-%m-%d UTC"
  )
}

mur_image_row <- function(times) {
  tags$div(
    class = "mur-sst-row",
    lapply(times, function(time_iso) {
      tags$div(
        class = "mur-sst-panel",
        tags$img(
          src = mur_image_url(time_iso),
          alt = paste("SST anomaly on", format_time_label(time_iso))
        ),
        tags$p(format_time_label(time_iso))
      )
    })
  )
}

mur_image_animator <- function(times, interval_ms = 900) {
  frame_urls <- vapply(times, mur_image_url, character(1))
  frame_labels <- vapply(times, format_time_label, character(1))
  widget_id <- "mur-sst-anim"

  tags$div(
    class = "mur-sst-anim",
    id = widget_id,
    `data-frame-urls` = as.character(jsonlite::toJSON(frame_urls, auto_unbox = TRUE)),
    `data-frame-labels` = as.character(jsonlite::toJSON(frame_labels, auto_unbox = TRUE)),
    `data-interval-ms` = interval_ms,
    tags$div(
      class = "mur-sst-anim-frame",
      tags$img(
        id = paste0(widget_id, "-img"),
        src = frame_urls[[1]],
        alt = paste("SST anomaly animation starting", frame_labels[[1]])
      )
    ),
    tags$div(
      class = "mur-sst-anim-controls",
      tags$button("Prev", type = "button", class = "btn btn-default mur-sst-prev"),
      tags$button("Pause", type = "button", class = "btn btn-primary mur-sst-toggle"),
      tags$button("Next", type = "button", class = "btn btn-default mur-sst-next"),
      tags$span(
        class = "mur-sst-anim-label",
        id = paste0(widget_id, "-label"),
        frame_labels[[1]]
      )
    ),
    tags$script(HTML("
(function() {
  function initMurSstAnim(root) {
    if (!root || root.dataset.murSstInit === 'true') return;
    root.dataset.murSstInit = 'true';

    const urls = JSON.parse(root.dataset.frameUrls);
    const labels = JSON.parse(root.dataset.frameLabels);
    const intervalMs = Number(root.dataset.intervalMs || 900);
    const img = root.querySelector('img');
    const label = root.querySelector('.mur-sst-anim-label');
    const toggleBtn = root.querySelector('.mur-sst-toggle');
    const prevBtn = root.querySelector('.mur-sst-prev');
    const nextBtn = root.querySelector('.mur-sst-next');

    let index = 0;
    let playing = true;
    let timer = null;

    function showFrame(nextIndex) {
      index = (nextIndex + urls.length) % urls.length;
      img.src = urls[index];
      img.alt = 'SST anomaly on ' + labels[index];
      label.textContent = labels[index];
    }

    function setPlaying(state) {
      playing = state;
      toggleBtn.textContent = playing ? 'Pause' : 'Play';
      if (playing) {
        timer = window.setInterval(function() {
          showFrame(index + 1);
        }, intervalMs);
      } else if (timer !== null) {
        window.clearInterval(timer);
        timer = null;
      }
    }

    toggleBtn.addEventListener('click', function() {
      setPlaying(!playing);
    });
    prevBtn.addEventListener('click', function() {
      showFrame(index - 1);
    });
    nextBtn.addEventListener('click', function() {
      showFrame(index + 1);
    });

    setPlaying(true);
  }

  document.querySelectorAll('.mur-sst-anim').forEach(initMurSstAnim);
})();
"))
  )
}

available_times <- fetch_times()
latest_times <- tail(available_times, n_days)
latest_time <- tail(available_times, 1)
```

## Latest daily maps

The row below shows the most recent `r n_days` daily SST anomaly maps for the Florida Keys region
(`r lat_min`°–`r lat_max`°N, `r lon_min`°–`r lon_max`°E), pulled directly from
[ERDDAP `r dataset_id`](`r erddap_base`.graph).

```{r image-row, results='asis'}
htmltools::tagList(mur_image_row(latest_times))
```

## Animation

Use the controls to pause auto-play or step backward and forward through the same daily maps.

```{r image-animation, results='asis'}
htmltools::tagList(mur_image_animator(latest_times))
```

**Map extent:** `r lat_min`°N to `r lat_max`°N, `r lon_min`°W to `r lon_max`°W.

**Color scale:** SST anomaly (`sstAnom`) from −3°C to +3°C (blue = cooler than climatology, red = warmer).

**Most recent analysis time in ERDDAP:** `r format_time_label(latest_time)`.

## Dataset details

| | |
|---|---|
| **Title** | Multi-scale Ultra-high Resolution (MUR) SST Analysis Anomaly fv04.1, Global, 0.01°, 2002–present, Daily |
| **Dataset ID** | [`jplMURSST41anom1day`](https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41anom1day.html) |
| **Institution** | NOAA NMFS SWFSC ERD and NOAA NESDIS CoastWatch WCRN |
| **Source product** | [JPL MUR L4 GHRSST v4.1](https://podaac.jpl.nasa.gov/dataset/MUR-JPL-L4-GLOB-v4.1) foundation SST |
| **Variable** | `sstAnom` — sea surface foundation temperature anomaly (°C) |
| **Climatology** | JPL daily MUR SST climatology, 2003–2014 |
| **Spatial resolution** | 0.01° (~1 km) |
| **Temporal resolution** | Daily (analysis time 09:00 UTC) |
| **Coverage** | Global, 2002-06-01 to present |
| **Platforms / sensors** | Aqua, Terra, Suomi-NPP, DMSP, NOAA-POES; AMSR-E, MODIS, VIIRS, AVHRR, SSM/I, in-situ |
| **License** | Free to use and redistribute; not intended for legal use ([full disclaimer](https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41anom1day.html)) |

### How the anomaly is computed

ERDDAP subtracts the JPL MUR 2003–2014 daily climatology from each day's foundation SST analysis.
For leap years, day-of-year 366 uses the day 365 climatology value.
The most recent four days are typically revised each day; older days may also be updated.

### ERDDAP image URL pattern

Each panel above is generated with a `griddap` PNG request of the form:

```text
https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41anom1day.smallPng
  ?sstAnom[(TIME)][(24.3):(26.5)][(-83.0):(-79.5)]
  &.draw=surface
  &.vars=longitude|latitude|sstAnom
  &.colorBar=|RedBlue| | |-3|3|
  &.land=over
```

Use the [Make A Graph](https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41anom1day.graph) or
[Data Access Form](https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41anom1day.html) pages to
explore other extents, color bars, or file formats.

### Related links

- [Dataset metadata](https://coastwatch.pfeg.noaa.gov/erddap/info/jplMURSST41anom1day/index.html)
- [PO.DAAC background](https://podaac.jpl.nasa.gov/dataset/MUR-JPL-L4-GLOB-v4.1)
- [ERDDAP griddap documentation](https://coastwatch.pfeg.noaa.gov/erddap/griddap/documentation.html)
