---
title: "NDBC Salinity Map"
subtitle: "Most recent readings from Everglades National Park buoys via SECOORA ERDDAP"
format: html
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(httr2)
library(htmltools)
library(knitr)
library(leaflet)
library(readr)
library(scales)
library(viridisLite)

# Station list and parameters mirror ndbc_fetch_airflow_dag.py
inwater_params <- c("sea_water_practical_salinity", "sea_water_temperature")

ndbc_stations <- list(
  list(name = "bkyf1", long_name = "Buoy_Key"),
  list(name = "bnkf1", long_name = "Butternut_Kay"),
  list(name = "bobf1", long_name = "Bob_Allen_Key"),
  list(name = "bwsf1", long_name = "Blackwater_Sound"),
  list(name = "dkkf1", long_name = "Duck_Key"),
  list(name = "gbtf1", long_name = "Garfield_Bight"),
  list(name = "hcef1", long_name = "Highway_Creek"),
  list(name = "jbyf1", long_name = "Joe_Bay"),
  list(name = "jkyf1", long_name = "Johnson_Key"),
  list(name = "lbsf1", long_name = "Little_Blackwater_Sound"),
  list(name = "lmdf1", long_name = "Little_Madeira_Bay"),
  list(name = "lrkf1", long_name = "Little_Rabbit_Key"),
  list(name = "lsnf1", long_name = "Long_Sound"),
  list(name = "mdkf1", long_name = "Middle_Key"),
  list(name = "mukf1", long_name = "Murray_Key"),
  list(name = "pkyf1", long_name = "Peterson_Key"),
  list(name = "tbyf1", long_name = "Terrapin_Bay"),
  list(name = "thrf1", long_name = "Thursday_Point"),
  list(name = "wrbf1", long_name = "Whipray_Basin")
)

# 25-hour lookback window
end_time <- format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
start_time <- format(Sys.time() - 25 * 3600, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")

fetch_latest_salinity <- function(station_info) {
  param <- "sea_water_practical_salinity"
  url <- sprintf(
    "https://erddap.secoora.org/erddap/tabledap/gov-nps-ever-%s.csv?time%%2Clatitude%%2Clongitude%%2C%s&time%%3E=%s&time%%3C=%s",
    station_info$name,
    param,
    start_time,
    end_time
  )

  resp <- request(url) |>
    req_perform()

  df <- read_csv(resp_body_string(resp), show_col_types = FALSE)
  names(df) <- c("time", "latitude", "longitude", "salinity")
  df <- df[-1, ] |>
    mutate(
      time = as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
      latitude = as.numeric(latitude),
      longitude = as.numeric(longitude),
      salinity = as.numeric(salinity)
    ) |>
    filter(!is.na(salinity))

  if (nrow(df) == 0) {
    return(NULL)
  }

  latest <- df[which.max(df$time), ]
  tibble(
    station_id = station_info$name,
    station = station_info$long_name,
    location = sprintf(
      "%s (%s)",
      station_info$long_name,
      sub("f1$", "", station_info$name)
    ),
    time = latest$time,
    latitude = latest$latitude,
    longitude = latest$longitude,
    salinity = latest$salinity
  )
}

salinity_data <- bind_rows(lapply(ndbc_stations, fetch_latest_salinity))
report_time <- format(Sys.time(), "%Y-%m-%d %H:%M %Z")
```

Data are pulled from [SECOORA ERDDAP](https://erddap.secoora.org/erddap/index.html):
using a 25-hour window ending at render time, with the most recent non-missing salinity value kept for each station.

**Report generated:** `r report_time`  
**Query window:** `r start_time` to `r end_time`

```{r summary}
if (nrow(salinity_data) == 0) {
  stop("No salinity data returned from ERDDAP for the current time window.")
}

salinity_data |>
  arrange(desc(time)) |>
  mutate(
    time = format(time, "%Y-%m-%d %H:%M UTC"),
    salinity = number(salinity, accuracy = 0.1)
  ) |>
  select(Station = station, `Station ID` = station_id, Time = time, `Salinity (PSU)` = salinity) |>
  kable()
```

```{r map}
#| fig-width: 10
#| fig-height: 7

salinity_palette <- colorNumeric(
  palette = c("#2166ac", "#67a9cf", "#d1e5f0", "#fddbc7", "#ef8a62", "#b2182b"),
  domain = salinity_data$salinity,
  na.color = "#999999"
)

leaflet(salinity_data) |>
  addProviderTiles("CartoDB.Positron") |>
  setView(
    lng = mean(salinity_data$longitude),
    lat = mean(salinity_data$latitude),
    zoom = 9
  ) |>
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 8,
    stroke = TRUE,
    weight = 1,
    color = "#333333",
    fillColor = ~salinity_palette(salinity),
    fillOpacity = 0.85,
    label = ~sprintf(
      "%s: %.1f PSU\n%s",
      location,
      salinity,
      format(time, "%Y-%m-%d %H:%M UTC")
    )
  ) |>
  addLegend(
    position = "bottomright",
    pal = salinity_palette,
    values = salinity_data$salinity,
    title = "Salinity (PSU)",
    opacity = 0.9
  )
```

Salinity is shown in practical salinity units (PSU, parts per thousand). Lower values (blue) indicate fresher water; higher values (red) indicate more marine conditions.

## Interpolated salinity field

The field below uses **DIVAnd** (Data-Interpolating Variational Analysis) with bathymetry masking, following the same approach as [sfer-mbon-oxygen](https://github.com/usf-imars/sfer-mbon-oxygen): ~0.01° horizontal grid resolution (~1 km), a two-scale analysis (smooth `0.08`° background + `0.035`° local anomalies with `epsilon2 = 0.01`), and interpolation run separately within each connected sea component so distant buoys do not smooth over local features. Land is masked using Natural Earth 10 m coastlines plus NOAA BlueTopo **4 m** bathymetry (8 m tiles fill gaps); cells with missing bathymetry or elevation ≥ −0.2 m are treated as land. Sea cells are extended along water paths from each buoy (not across islands), then short bridges are added only through BlueTopo water cells (elevation < −0.2 m) to connect nearby stations still isolated from the main field.

```{r interpolate-field}
obs_file <- file.path("data", "interpolated", "salinity_obs.csv")
field_file <- file.path("data", "interpolated", "salinity_field.csv")
dir.create(dirname(obs_file), recursive = TRUE, showWarnings = FALSE)

salinity_data |>
  select(longitude, latitude, salinity) |>
  write_csv(obs_file)

interp_log <- system2(
  "scripts/run_interpolate_field.sh",
  c(obs_file, field_file, "ndbc_buoys"),
  stdout = TRUE,
  stderr = TRUE
)
exit_code <- attr(interp_log, "status")
if (!is.null(exit_code) && exit_code != 0L) {
  stop(paste("Julia interpolation failed:\n", paste(interp_log, collapse = "\n")))
}
cat(paste(interp_log, collapse = "\n"), "\n")

salinity_field <- read_csv(field_file, show_col_types = FALSE)
cat("Interpolated field cells:", nrow(salinity_field), "\n")
source("R/plot_ndbc_field.R")
```

```{r interpolated-map}
#| fig-width: 10
#| fig-height: 7

if (nrow(salinity_field) == 0) {
  stop("Interpolated salinity field is empty.")
}

plot_salinity_field(salinity_field, salinity_data)
```

The colored grid shows the DIVAnd-interpolated field; circle markers show the buoy observations used as input.
