---
title: "NDBC Salinity Gauges"
subtitle: "Latest buoy salinity with low / normal / high thresholds"
format: html
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(httr2)
library(htmltools)
library(readr)
library(scales)

source("R/plot_salinity_gauges.R")

ndbc_stations <- list(
  list(name = "bkyf1", long_name = "Buoy_Key"),
  list(name = "bnkf1", long_name = "Butternut_Kay"),
  list(name = "bobf1", long_name = "Bob_Allen_Key"),
  list(name = "bwsf1", long_name = "Blackwater_Sound"),
  list(name = "dkkf1", long_name = "Duck_Key"),
  list(name = "gbtf1", long_name = "Garfield_Bight"),
  list(name = "hcef1", long_name = "Highway_Creek"),
  list(name = "jbyf1", long_name = "Joe_Bay"),
  list(name = "jkyf1", long_name = "Johnson_Key"),
  list(name = "lbsf1", long_name = "Little_Blackwater_Sound"),
  list(name = "lmdf1", long_name = "Little_Madeira_Bay"),
  list(name = "lrkf1", long_name = "Little_Rabbit_Key"),
  list(name = "lsnf1", long_name = "Long_Sound"),
  list(name = "mdkf1", long_name = "Middle_Key"),
  list(name = "mukf1", long_name = "Murray_Key"),
  list(name = "pkyf1", long_name = "Peterson_Key"),
  list(name = "tbyf1", long_name = "Terrapin_Bay"),
  list(name = "thrf1", long_name = "Thursday_Point"),
  list(name = "wrbf1", long_name = "Whipray_Basin")
)

end_time <- format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
start_time <- format(Sys.time() - 25 * 3600, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")

fetch_latest_salinity <- function(station_info) {
  param <- "sea_water_practical_salinity"
  url <- sprintf(
    "https://erddap.secoora.org/erddap/tabledap/gov-nps-ever-%s.csv?time%%2Clatitude%%2Clongitude%%2C%s&time%%3E=%s&time%%3C=%s",
    station_info$name,
    param,
    start_time,
    end_time
  )

  resp <- request(url) |>
    req_perform()

  df <- read_csv(resp_body_string(resp), show_col_types = FALSE)
  names(df) <- c("time", "latitude", "longitude", "salinity")
  df <- df[-1, ] |>
    mutate(
      time = as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
      latitude = as.numeric(latitude),
      longitude = as.numeric(longitude),
      salinity = as.numeric(salinity)
    ) |>
    filter(!is.na(salinity))

  if (nrow(df) == 0) {
    return(tibble(
      station_id = station_info$name,
      station = station_info$long_name,
      location = sprintf(
        "%s (%s)",
        station_info$long_name,
        sub("f1$", "", station_info$name)
      ),
      time = as.POSIXct(NA, origin = "1970-01-01", tz = "UTC"),
      latitude = NA_real_,
      longitude = NA_real_,
      salinity = NA_real_
    ))
  }

  latest <- df[which.max(df$time), ]
  tibble(
    station_id = station_info$name,
    station = station_info$long_name,
    location = sprintf(
      "%s (%s)",
      station_info$long_name,
      sub("f1$", "", station_info$name)
    ),
    time = latest$time,
    latitude = latest$latitude,
    longitude = latest$longitude,
    salinity = latest$salinity
  )
}

salinity_data <- bind_rows(lapply(ndbc_stations, fetch_latest_salinity)) |>
  arrange(station)

report_time <- format(Sys.time(), "%Y-%m-%d %H:%M %Z")
n_low <- sum(salinity_data$salinity < SALINITY_LOW_THRESHOLD, na.rm = TRUE)
n_high <- sum(salinity_data$salinity > SALINITY_HIGH_THRESHOLD, na.rm = TRUE)
n_normal <- sum(
  !is.na(salinity_data$salinity) &
    salinity_data$salinity >= SALINITY_LOW_THRESHOLD &
    salinity_data$salinity <= SALINITY_HIGH_THRESHOLD
)
n_missing <- sum(is.na(salinity_data$salinity))
```

Data are pulled from [SECOORA ERDDAP](https://erddap.secoora.org/erddap/index.html) using the same 25-hour window as the salinity map. Each gauge shows the latest non-missing reading for an Everglades National Park buoy.

**Report generated:** `r report_time`  
**Query window:** `r start_time` to `r end_time`

## Thresholds

Gauges use practical salinity units (PSU) on a `r SALINITY_GAUGE_MIN`–`r SALINITY_GAUGE_MAX` scale suited to Florida Bay and the Keys:

```{r legend, results='asis'}
htmltools::tagList(salinity_gauge_legend())
```

- **Low** (`< `r SALINITY_LOW_THRESHOLD` PSU): fresher conditions, stronger freshwater or runoff influence
- **Normal** (`r SALINITY_LOW_THRESHOLD`–`r SALINITY_HIGH_THRESHOLD` PSU): typical estuarine range for the region
- **High** (`> `r SALINITY_HIGH_THRESHOLD` PSU): more marine or hypersaline conditions

**Current summary:** `r n_low` low, `r n_normal` normal, `r n_high` high, `r n_missing` missing.

## Station gauges

```{r gauges, results='asis'}
if (all(is.na(salinity_data$salinity))) {
  stop("No salinity data returned from ERDDAP for the current time window.")
}

htmltools::tagList(plot_salinity_gauges(salinity_data))
```

Thresholds are intended for quick situational awareness across stations. They are not regulatory standards and can be adjusted in `R/plot_salinity_gauges.R` if monitoring goals change.
