---
title: "NDBC Salinity Time Series"
subtitle: "Per-buoy salinity from the USF MBON Grafana dashboard"
format: html
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(ggplot2)
library(htmltools)
library(httr2)

source("R/plot_grafana_salinity_ts.R")

ndbc_stations <- list(
  list(name = "bkyf1", long_name = "Buoy_Key"),
  list(name = "bnkf1", long_name = "Butternut_Kay"),
  list(name = "bobf1", long_name = "Bob_Allen_Key"),
  list(name = "bwsf1", long_name = "Blackwater_Sound"),
  list(name = "dkkf1", long_name = "Duck_Key"),
  list(name = "gbtf1", long_name = "Garfield_Bight"),
  list(name = "hcef1", long_name = "Highway_Creek"),
  list(name = "jbyf1", long_name = "Joe_Bay"),
  list(name = "jkyf1", long_name = "Johnson_Key"),
  list(name = "lbsf1", long_name = "Little_Blackwater_Sound"),
  list(name = "lmdf1", long_name = "Little_Madeira_Bay"),
  list(name = "lrkf1", long_name = "Little_Rabbit_Key"),
  list(name = "lsnf1", long_name = "Long_Sound"),
  list(name = "mdkf1", long_name = "Middle_Key"),
  list(name = "mukf1", long_name = "Murray_Key"),
  list(name = "pkyf1", long_name = "Peterson_Key"),
  list(name = "tbyf1", long_name = "Terrapin_Bay"),
  list(name = "thrf1", long_name = "Thursday_Point"),
  list(name = "wrbf1", long_name = "Whipray_Basin")
)

from_ms <- GRAFANA_DEFAULT_FROM
to_ms <- GRAFANA_DEFAULT_TO

salinity_ts <- fetch_grafana_salinity_timeseries(
  from = from_ms,
  to = to_ms
)

report_time <- format(Sys.time(), "%Y-%m-%d %H:%M %Z")
range_start <- format(
  as.POSIXct(as.numeric(from_ms) / 1000, origin = "1970-01-01", tz = "UTC"),
  "%Y-%m-%d"
)
range_end <- format(
  as.POSIXct(as.numeric(to_ms) / 1000, origin = "1970-01-01", tz = "UTC"),
  "%Y-%m-%d"
)
```

Each panel below matches **Grafana panel 33** (*In-situ salinity (psu) at ~1m depth from $Buoy*) on the [USF MBON South Florida Salinity dashboard](`r grafana_dashboard_url(from_ms, to_ms)`). Time series are queried from the dashboard's InfluxDB datasource via the Grafana `/api/ds/query` endpoint, using the same query and time range as the dashboard link you provided.

**Report generated:** `r report_time`  
**Time range:** `r range_start` to `r range_end` UTC  
**Stations plotted:** `r length(unique(salinity_ts$location))`

Dashed lines mark the same low (`r 25` PSU) and high (`r 35` PSU) thresholds used on the [Salinity Gauges](ndbc_salinity_gauges.qmd) page.

## Buoy panels

```{r buoy-panels, fig.width=9, fig.height=3.2, results='asis'}
if (nrow(salinity_ts) == 0) {
  stop("No salinity time series returned from the Grafana datasource.")
}

htmltools::tagList(
  plot_grafana_salinity_panels(
    ndbc_stations,
    salinity_ts,
    from = from_ms,
    to = to_ms
  )
)
```

## Grafana links

- [Full South Florida Salinity dashboard](`r grafana_dashboard_url(from_ms, to_ms)`)
- Panel ID: `r GRAFANA_SALINITY_PANEL_ID` on dashboard `r GRAFANA_DASHBOARD_UID`

Grafana solo-panel iframes are not embedded here because the MBON server sets `X-Frame-Options: deny` and does not expose the image renderer plugin. Use the **Open Grafana panel** link on any chart to view the live Grafana panel for that buoy.
